An Expressive and Efficient Solution to the Service Selection Problem

### 2010-06-29 publicaciones/iswc10/paperISWC10-v3.tex
paperISWC10-v3

% Services
\newcommand{\nationalFlight}{\textit{national-flight}}
\newcommand{\onewayFlight}{\textit{one-way-flight}}
\newcommand{\nationalTrain}{\textit{national-train}}
\newcommand{\onewayTrain}{\textit{one-way-train}}
\newcommand{\onestop}{\textit{one-stop}}
\newcommand{\toPA}{\textit{to-pa}}
\newcommand{\onestopPA}{\textit{onestop-to-pa}}
\newcommand{\fromNY}{\textit{from-ny}}
\newcommand{\fromLA}{\textit{from-la}}

@@

\begin{abstract}
Given the tremendous amount of potential Semantic Web 
Services that can be created from online sources by using existing 
annotation tools, expressive formalisms and efficient and scalable approaches to solve the Service Selection Problem (SSP) are required. In this paper, we propose a framework that 
is firmly grounded on logic. The framework adopts the Local-As-View 
approach~\cite{levy:bucket} that semantically describes services in terms of concepts in a domain 
ontology. A user request is comprised of a conjunctive query on concepts 
from the domain ontology and a set of preferences; the query expresses 
the abstract concepts that must be combined to solve the query, and 
the preferences are used to rank the solutions. Relations in the domain 
ontology are considered to augment the space of available services. Using 
this representation, the problem of service selection is cast as a problem 
of query rewriting, from the area of integration systems. Then, 
building on related work, we devise an encoding of SSP as a logical theory 
whose models are in correspondence with the service rewritings of the 
user query, and in presence of preferences, whose best models are in 
correspondence with the best-ranked valid solutions. Thus, by exploiting 
known properties of modern SAT solvers, we provide an efficient and 
scalable solution to SSP. The approach provides the basis to represent 
a large number of real-world situations; additionally, it does not 
only scale up to large instances, as the experimental results show, 
but is also sound and complete since, being founded on logic, is 
amenable to formal analysis.
\end{abstract}

\section{Introduction}
Existing Web infrastructures support the publication and access to a
tremendous amount of Web data sources, some of which can be labeled
and converted into  Semantic Web Services by using existing annotation
tools like the one proposed  by Ambite et al. \
\cite{AmbiteISWC09}. Once this huge dataset of Semantic Web  Services
become available, users more than ever will require techniques to
effectively select the services that meet their requirements. In order
to achieve  this goal, functional and non-functional properties of the
services should be  precisely described, as well as criteria to select
and rank the solutions that best meet user requests.   In this paper,
we extend an approach traditionally used in the area of data
integration to solve the Service Selection Problem. First, concepts
from a domain ontology provide a precise description of the semantics
of available sources. These  descriptions correspond to views on
concepts of the domain ontology and follow  the Local-As-View approach
(LAV)~\cite{levy:bucket}. LAV mappings are able to  scale to service
datasets that constantly change, in contrast to Global-As-View
definitions  (GAV), where generic concepts are defined in terms of the
available sources and  cannot be easily modified when sources
constantly change. A user request is  modeled as a conjunctive query
on the ontology concepts, and user preferences  are associated with a
cost and constraint the solutions of the query.  Based on this
integration framework, we define the problem of selecting the
services that best meet the user request, as the problem of
enumerating and  ranking the service rewritings of the user
query. Each rewriting is associated  with a cost equal to the cost of
the preferences violated by the rewriting, and the best ranked
rewriting is one that has minimum cost among all valid
rewritings. Knowledge encoded in the domain ontology is used to expand
the  space of rewritings. Thus, our proposed service selection
approach is cast into  the well-known Query Rewriting Problem (QRP)
for LAV which is central to  integration systems.  The QRP  has been shown to
be NP-complete~\cite{Ullman00}, and it  consists of a conjunctive
query that must be answered in terms of views in  which the query and
the views are described using LAV with abstract relations.  This
problem is important in the context of data
integration~\cite{Chen05,JaudoinPRST05},  and query optimization and
data maintenance~\cite{AfratiLU07,levy:bucket}, and  several
approaches have been defined that scale up to a large number of
views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.

We propose a solution that extends the recent approach of Arvelo et
al.\ \cite{arvelo:aaai06} which is based on the efficient enumeration
of models of a propositional logic  theory. Similarly, given an
instance of SSP, a logical theory is constructed such  that each model
of the theory encodes a valid rewriting; the theory can be constructed
in polynomial time from the instance of the SSP. All valid
rewritings  are obtained by enumerating all models of the theory in
linear time by using  off-the-shelf model enumeration solvers such as
c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-shelf SAT solvers
such as Minisat~\cite{minisat} can be used to produce all the minimal
rewritings. Thus, we exploit known properties of modern SAT solvers
that  implement non-chronological backtracking via clause learning,
caching of common  subproblems and heuristics, to provide an efficient
and scalable solution to SSP.

To summarize the contributions of our work are as follows:
\begin{itemize}
\item Service descriptions as LAV mappings on ontology concepts and user preferences as logic constraints, to provide an expressive description framework.
\item Formalization of SSP as the well-known QRP, and the extension of efficient and scalable existing solutions to exploit the properties of modern SAT solvers to construct a logical theory that encodes an instance of SSP, in a way that the theory models encode the valid rewritings and best models 
correspond to the optimal (best) rewritings.
\item Reduction of the combinatorial search, by efficiently enumerating the solutions.
\end{itemize}

The paper contains five more sections. The next section defines our
approach  and illustrates the SSP with typical examples. Section 3
describes our solution  and system architecture; Section 4 reports our
empirical results over different benchmark problems. Sections 5
summarizes existing approaches; and the paper  concludes with a final
discussion in Section 6.

\section{Service Selection Problem}
We consider service selection problems comprised of the
description $IS$ of the integration framework and of a
user request $R$ over the $IS$.

Our setting is based on the assumption that services are
semantically described in terms of a domain ontology. 
Formally, the integration framework corresponds to a tuple
$IS=\tup{D,S,M}$ where $D$ is the domain ontology, $S$ describes
the available services, and $M$ is the collection of LAV mappings
that define the services in $S$ as views on the ontology $D$. 
On the other hand, the user input $R=\tup{Q,P}$ consists of a
request for composition of services, described as a conjunctive
query $Q$ over the ontology, and a set $P$ of preferences for
the requested composition.

In the following subsections, we describe in detail all the
ingredients making up a service selection problem, and illustrate
the framework with several examples.

@@

The domain ontology $D$ is a tuple $\tup{\sigma,A}$ where $\sigma$ is a
\emph{signature} for a logical language and $A$ is a collection of axioms
describing the ontology.
A signature $\sigma$ is a tuple $\tup{R_1^{r_1},\ldots,R_n^{r_n},c_1,\ldots,c_m}$
where each $R_i$ is a relational symbol of arity $r_i$,\footnote{We use
the notation $R^r$ to say that $R$ is a relational symbol of arity $r$.}
and each $c_j$ is a constant symbol.
The set of axioms describe the ontology by defining the relationships
between the ontology relations. % Not sure if talk about Datalog here...
For the present work, we only assume subsumption relationships among
concepts that are expressed with rules of the form:
\begin{equation}
\label{eq:orule}
R(\bar x,\bar y, \bar a)\ \orule\ P(\bar x, \bar b)\,,
\end{equation}
where $R$ and $P$ are predicates in $\sigma$ of appropriate arity,
$\bar x$ and $\bar y$ are lists of variables (repetitions allowed),
and $\bar a$ and $\bar b$ are lists of constant symbols in $\sigma$
(repetitions allowed). All these lists except $\bar x$ can be empty.

Although limited in appearance, rules of this form are quite expressive
as they allow us to specify diverse relationships among predicates.
For example,
\begin{itemize}
\item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
classes are specified with unary predicates. A subclass relationship
among two classes is directly specified with a simple rule; e.g., the
rule $\textit{penguin}(x) \orule \textit{bird}(x)$ declares that every
penguin is a bird.
\item \textbf{Subrelations via specialization:} A subrelation of $R^r$
can be specified by constraining another relation $P^s$ ($r<s$). 
For example, the rule:
\begin{equation}
\label{eq:QE2}
\textit{descendant}(\textit{Elizabeth II},x)\ \orule\ \textit{noble}(x)
\end{equation}
tells that any descendant of Queen Elizabeth II is a noble.
\item \textbf{Indirect subsumption:} it is even possible to specify
a subrelation via another seemingly unrelated predicate. For example,
the rule:
\[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
says that when the second argument of `CitizenOf' is fixed to the
constant `Montreal', the tuples in the relation `CitizenOf' are
contained in the set of tuples in the relation `LivesIn' whose
second component is `Canada'.
\end{itemize}
However, we require that the \emph{dependency graph} for the ontology
to be a \emph{forest of trees}. The dependency graph $G(D)$ is a
labeled directed graph that is constructed as follows: the nodes
are the relational symbols in the signature $\sigma$, and there
is an edge $(R,P)$ in the graph iff there is an rule of the
form: \eqref{eq:orule} that is labeled with the \emph{bindings}
induced by the rule. For example, if $\textit{descendant}(x,y)$
and $\textit{noble}(z)$ are two predicates in the signature
and there is the rule \eqref{eq:QE2}, there edge from 
Descendant to Noble is labeled with the bindings
$\{x=\textit{Elizabeth II},y=z\}$.

\subsection{Services and LAV Mappings}

The available services are represented by means of another signature
$S=\tau=\tup{S_1^{s_1},\ldots,S_k^{s_k}}$, called the \emph{services signature},
that consists of relational symbols $S_i$ of arity $s_i$: each such symbol
$S_i$ represents a concrete service in the Semantic Web that ``offers''
some information.

The semantic description of services is expressed following the
LAV paradigm in terms of LAV mappings. These mappings describe the
services in terms of concepts in the domain ontology \cite{Ullman00}:
for each service $S_i$, there is a LAV mapping that describes $S_i$
as a \emph{conjunctive query} on the concepts in the ontology; 
input and output attributes of the service are also distinguished.
For example, a service $S(x,y)$ that returns information about
flights originating at a given US city can be described with the view:
\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
where $flight^2$ and $uscity^1$ are relational symbols in the
ontology; the symbol \$ denotes that $x$ is an input attribute. 
The semantic interpretation of a mapping like this one enforce the following:
\begin{itemize}
\item the service represented by $S$ provides information in the
      form of tuples $(x,y)$,
\item each tuple $(x,y)$ returned by the service satisfies the rhs
      of the view; i.e., $\flight(x,y)$ and $\UScity(x)$, and
\item the views are not necessarily \emph{complete}; i.e., there may be
      other tuples $(x,y)$ that satisfy the rhs of the view but which
      are not available through the service represented by $S$.
\end{itemize}

The LAV approach is commonly used in integration systems because it permits
the scalability of the system as new services become available \cite{Ullman00}.
Under LAV, the appearance of a new service only causes the addition of a new
mapping describing the service in terms of the concepts in the ontology.
Under GAV, on the other hand, the ontology concepts are semantically described
using views in terms of the sources of information.
Thus, the extension or modification of the ontology is an easy task in GAV as
it only involves the addition or local modification of few descriptions
\cite{Ullman00}.

Therefore, the LAV approach is best suited for applications with a stable
ontology but with changing data sources whereas the GAV approach is best
suited for applications with stable data sources and a changing ontology.
In any case, the chosen approach is guarantee to efficiently scale with
respect to the changing characteristic but not with respect to the stable one;
e.g., a LAV description can be efficiently updated as data sources
are added or deleted, yet a slight revision of the ontology implies a
revision of all LAV mappings.

For the Semantic Web, we assume that the ontology of concepts is the stable
component. We believe that this is a reasonable assumption since, once a
common language is agreed upon to describe Web resources, the only changing
characteristic is the number and nature of resources which constantly pop
up or disappear from the Web.

Up to here, we have described all elements in the integration framework
$IS=\tup{D,S,M}$ where $D=\tup{\sigma,A}$ is an ontology of concepts, 
$S=\tau$ represent the available services in the Web and $M$ is a
collection of LAV mappings that describe the services in terms of the
concepts in the ontology.
The integration framework can be thought as the ``knowledge base'' (KB)
in a system designed for answering requests about the selection and
composition of Web services.
Ideally, the KB should support the efficient processing of user requests.

@@

A user request is formally described by a tuple $R=\tup{Q,P}$
where $Q$ is a conjunctive query in terms of concepts from the
ontology that expresses how the abstract concepts must be 
combined to resolve a given task, and a set $P$ of preferences.
For example, the query:
\[ Q(x)\ \qrule\ \flight(\LA,x), \flight(x,\PA) \]
can be used to find all cities on which a two-leg flight
from Los Angeles to Paris stop. Continuing with the above example,
this query can be answered using the view represented by $S$
as $I(x) \qrule S(\LA,x), S(x,\PA)$. This rewriting is correct yet
not necessarily complete as there may be two-leg flights from
Los Angeles to Paris that stop at non-US cities, e.g.,\ London.

The preferences are used to rank the collection of all valid
rewritings of the query. Once this ranking is obtained, the solution
of request $R$ is any best-ranked valid rewriting.
In this work, we consider a simple yet expressive system of
preferences in which each preference is given as a 
`soft constraint' on the rewriting of the query.
A soft constraint is a tuple $\pi=\tup{\varphi,c}$ where $\varphi$
is a propositional formula and $c$ is the cost associated with $\varphi$.
The idea is that each valid rewriting of the query is associated with
a cost equal to the sum of the costs of the preferences violated by
the rewriting, and that the best ranked rewriting is one that has minimum
cost among all valid rewriting.

It only remains to say over which language the formulas $\varphi$
are allowed to be and when a preference is violated by a rewriting.
The set of propositions for constructing preferences is
$\L(IS)=\{ R : R\in\sigma \} \cup \{S : S\in\tau \}$ that corresponds
to the relational symbols either in the ontology signature $\sigma$
or in the services signature $\tau$. Elements of $\L(IS)$ are
propositional symbols that should not be confused with their
relational interpretation in $IS$; indeed, if the reader
is more comfortable, he may use a different symbols altogether
such as $P_R$, $[R]$, or other.

The validity of a preference is established with respect to the
propositional model $\M(I)$ (truth assignment for the symbols in $\L(IS)$)
constructed from an answer $I(\bar x)$.
The model $\M(I)$ is defined by $S=\textbf{true}$ iff the service
$S$ appears in $I(\bar x)$, for $S\in\tau$, or by $R=\textbf{true}$
iff the concept $R$ appears in the unique path from a concept 
$R'$ to the root in the dependency graph $G(D)$ where $R'$ appears
in a service $S(\bar y)$ used in $I(\bar x)$, for $R\in\sigma$.
That is, the model makes true the service symbols used in
$I(\bar x)$, or the ontology symbols used in services in $I(\bar x)$,
or the ontology symbols that can be reached from the latter in
the dependency graph $G(D)$.
For example, the answer $I(x)\qrule S(\LA,x),S(x,\PA)$ defines
the model $\M(I)=\{S=\textbf{true},\flight=\textbf{true},\UScity=\textbf{true}\}$.

A preference $\varphi$ holds in an answer $I(\bar x)$ iff
$\M(I)\vDash\varphi$. This simple system of preferences permits
us to represent interesting cases such as:
\begin{itemize}
\item \textbf{Hard constraints:} any constraint $\pi=\tup{\varphi,\infty}$
is treated as a hard constraint that must be satisfied by every rewriting,
\item \textbf{QoS preferences:} this type of preferences can be used to
assign absolute quantities of reward/cost to single services as the one
used for integrated QoS parameters. For example, if each service $S_i$
is associated with a QoS reward of $r_i$, then the collection of preferences
$\pi_i=\tup{\neg S_i,-r_i}$ selects a valid rewriting with services that
have the highest combined QoS as the best rewriting,
\item \textbf{Conditional preferences:} a user's preference of the type
`if service $S$ is used, then service $R$ should be used as well' can be
modeled with the hard constraint $S \Rightarrow R$,
\item \textbf{Preferences of the type at-least-one:} a user's preference
of the type that at least one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the hard constraint $S_1\lor\cdots\lor S_n$, and
\item \textbf{Preferences of the type at-most-one:} a user's preference
of the type that at most one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the collection
$\{\neg S_i\lor\neg S_j: i\neq j\}$ of hard constraints.
\end{itemize}

@@

We extend the \mcdsat\ system of Arvelo et al.\ \cite{arvelo:aaai06} for
Query Rewriting Problems (QRP). An instance of a QRP consists of a
collection of views and a query on abstract concepts. 
The problem consists is in ``re-writing'' the query in terms of the
views so that each tuple produced by the rewriting is a tuple of the
solution.
\mcdsat\ reduces the QRP into the problem of finding the models of
a propositional logic theory.
The theory constructed by \mcdsat\ satisfies the following properties:
(1)~there is a 1-1 correspondence between the valid rewritings of the
query and the models of the logical theory, (2)~given a model of the theory,
one can recover the corresponding rewriting in linear time, and
(3)~the theory can be constructed in polynomial time from the instance 
of the QRP.
Once the logical theory is constructed, one can be interested in finding
all minimal rewritings of the query as done in data integration systems
with incomplete sources \cite{Ullman00}, or just a model of the theory
\cite{refs}. For the former, off-the-shelf model enumeration solvers
such as c2d \cite{c2d} and Relsat \cite{relsat} can be used, while 
off-the-shelf SAT solvers such as Minisat \cite{minisat} can be used in the
latter case.

In this section, we have just enough space to explain how the logical
theory constructed by \mcdsat\ can be extended to capture the features
associated with SSPs that are not present in QRPs; namely, handling
constant symbols and input and output attributes, the ontology of
concepts with subsumption relationships, and user's preferences. The
result is an extended theory whose models are in correspondence with
the valid solutions of the SSP and, in the presence of preferences,
whose \emph{best models} are in correspondence with the best-ranked
valid solutions of the SSP.

\subsubsection{Constant Symbols and Input Output Attributes}

>>>>

% Services
\newcommand{\nationaltlight}{\textit{national-flight}}
\newcommand{\onewayflight}{\textit{one-way-flight}}
\newcommand{\nationaltrain}{\textit{national-train}}
\newcommand{\onewayTrain}{\textit{one-way-train}}
\newcommand{\onestop}{\textit{one-stop}}
\newcommand{\toPA}{\textit{to-pa}}
\newcommand{\onestopPA}{\textit{onestop-to-pa}}
\newcommand{\fromNY}{\textit{from-ny}}
\newcommand{\fromLA}{\textit{from-la}}

@@

\begin{abstract}
Given the large number of Semantic Web Services that can be created from
online sources by using existing annotation tools, expressive formalisms and
efficient and scalable approaches to solve the service selection problem are
required in order to make this services widely available to the users.
In this paper, we propose such a framework that is grounded on logic and
the Local-As-View approach for describing instances of the service selection problem.
In our approach, Web services are semantically described using LAV mappings
in terms of generic concepts from an ontology, while the user requests correspond
to conjunctive queries on the generic concepts, that must be combined to solve
a task, plus a set of preferences that are used to rank the possible solutions.
The LAV formulation allows us to cast the service selection problem as a
query rewriting problem that must consider the relationships among the concepts
in the ontology and the ranks induced by the preferences of the user.
Then, building on related work, we devise an encoding of the resulting
query rewriting problem as a logical theory whose models are in correspondence
with the solutions of the user request, and in presence of preferences, whose
best models are in correspondence with the best-ranked solutions.
Thus, by exploiting known properties of modern SAT solvers, we provide an
efficient and scalable solution to the service selection problem.
The approach provides the basis to represent a large number of real-world
situations and interesting user requests.
\end{abstract}

\section{Introduction}

Existing Web infrastructures support the publication and access to a
tremendous amount of Web data sources, some of which that can be labeled
and converted into Semantic Web Services by using existing annotation
tools like the one proposed by Ambite et al.\ \cite{AmbiteISWC09}.
Once a large dataset of Semantic Web Services become available, users
will require techniques to effectively select the services that meet
their requirements.
In order to achieve this goal, the services in the dataset must be
tagged with their functional and non-functional properties, and
the users' preferences and requirements must be formally described
as well.
In this paper, we extend an approach traditionally used in the area of
data integration to solve the problem of selecting the best services
that meet a user request, a problem that we call in this paper the
Service Selection Problem (SSP).

As in other approaches, we use domain ontologies for describing
the services in the dataset, yet we differ in how the services are described.
In this paper, we use the recent approach of Ambite et al.\ \cite{AmbiteISWC09}
that describes services as views on the generic concepts of the ontology
following the Local-As-View (LAV) approach that is widely used in integration
systems~\cite{levy:bucket}, instead of the traditional Global-As-View (GAV)
approach where the generic concepts are expressed in terms of the services.
The adoption of the LAV approach instead of GAV is not accidental.
LAV descriptions are tailored towards systems with constantly changing
datasets and a relatively stable set of generic concepts, while GAV
descriptions are tailored towards systems with a constantly changing
set of generic concepts but a relatively stable dataset of services.

As it is shown below, LAV descriptions are achieved with mappings
that define the services as conjunctive queries involving the generic
concepts in the ontology. Thus, every time that a service changes or
a new one becomes available, only a tiny fraction of the mappings
must be updated, usually just one mapping.
Likewise, user requests can be modeled as conjunctive queries over the
generic concepts in a way that the SSP can be cast as the problem
of rewriting a query in terms of a set of views, the so-called Query
Rewriting Problem (QRP) that is well-known in the area of data
integration~\cite{Chen05,JaudoinPRST05}, query optimization and data
maintenance~\cite{AfratiLU07,levy:bucket}, and for which several
scalable approaches had been
proposed~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.
Furthermore, users' preferences and constraints on the set of the
selected services to serve a given request are supported by the
adoption of a simple yet expressive language for preferences. 
These preferences and constraints further refine and rank the
the set of valid rewritings of the posed query, in a way that
the best solution to the SSP correspond to the best-ranked 
valid rewritings of the corresponding QRP.

Our solution for the SSP extends the recent approach of Arvelo et
al.~\cite{arvelo:aaai06} for QRPs that is based on the efficient 
enumeration of models for a propositional logic theory.
In our case, an input instance of the SSP is converted into an instance
of a QRP with preferences and constraints that is then translated
into a logical theory such that its best models are in correspondence
with the solutions of the SSP. These translations, from SSP to QRP to
logic, are performed efficiently, in (low) polynomial time, while the
best models are found using off-the-shelf SAT tools.
Thus, we are able to exploit the befits of modern SAT techniques such
as conflict-directed backtracking via clause learning and caching and
decomposition of common subproblems~\cite{beame:understanding} to
perform the necessary search in the combinatorial space of solutions.

In summary, we make the following crisp contributions to the problem of 
service selection and composition: (1)~propose the use of LAV mappings
as an scalable solution for describing the continuously changing set
of available Web Services in terms of stable generic concepts,
(2)~propose a simple yet powerful language for expressing preferences
and constraints on the valid solutions of the SSP, 
(3)~describe how to transform the SSP to the well-known QRP extended with
preferences and constraints, and
(4)~describe how to change an efficient and scalable solution to the QRP,
based on propositional logic and SAT tools, to handle preferences and constraints.
The rest of this paper is as follows. The next two sections describe the
SSP and the language of preferences and constraints, and the proposed solution to the SSP.
Then, we present preliminary experimental results, related work 
and end with a brief discussion.

\section{Service Selection Problem}

An SSP consists of a description $IS$ of the integration framework
and a user request $R$.
Formally, the integration framework is a tuple $IS=\tup{D,S,M}$
where $D$ is the ontology of generic concepts, $S$ is the set
of available services, and $M$ is the collection of LAV mappings that
semantically describe the services in terms of the ontology.
On the other hand, a user request is a tuple $R=\tup{Q,P}$
that is made of a query $Q$ expressed as a conjunctive query over
the generic concepts and a set of preferences $P$.
In the following, we describe all these elements in detail and
illustrate the framework through a number of examples.

@@

The domain ontology $D$ is a tuple $\tup{\sigma,A}$ where $\sigma$
is a \emph{signature} for a logical language and $A$ is a collection
of axioms describing the ontology.
A signature $\sigma$ is a set of relational and constant symbols from
which logical formulas can be constructed; it corresponds to a tuple
$\tup{R_1^{r_1},\ldots,R_n^{r_n},c_1,\ldots,c_m}$ where each $R_i$
is a relational symbol of arity $r_i$,\footnote{We use the notation
$R^r$ to say that $R$ is a relational symbol of arity $r$.} and each
$c_j$ is a constant symbol.
The axioms describe the ontology by defining the relationships between
the ontology concepts.
For the present work, we only consider subsumption relationships
between concepts that are expressed with rules of the form:
\begin{equation}
\label{eq:orule}
R(\bar x,\bar y, \bar a)\ \orule\ P(\bar x, \bar b)\,,
\end{equation}
where $R$ and $P$ are predicates in $\sigma$ (of appropriate arity),
$\bar x$ and $\bar y$ are lists of variables (repetitions allowed),
and $\bar a$ and $\bar b$ are lists of constant symbols (repetitions allowed).
All these lists may be empty except $\bar x\cup \bar b$.

Although limited in appearance, subsumption rules are quite expressive
as they allow us to specify diverse relationships between concepts; e.g.,
\begin{itemize}
\item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
classes are specified with unary predicates. A subclass relationship can
be specified with a simple rule; e.g.,
$\textit{penguin}(x) \orule \textit{bird}(x)$ tells that penguins are birds.
\item \textbf{Subrelations via specialization:} A subrelation of $R^r$
can be specified by constraining another relation $P^s$ ($r<s$). 
For example, the rule:
\begin{equation}
\label{eq:QE2}
\textit{descendant}(\textit{Elizabeth II},x)\ \orule\ \textit{noble}(x)
\end{equation}
tells that the descendants of Queen Elizabeth II are noble.
\item \textbf{Indirect subsumption:} it is even possible to specify
a subrelation via another seemingly unrelated predicate. For example,
the rule:
\[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
says that when the second argument of `\textit{citizen-of}' is fixed to the
constant `\textit{Montreal}', the tuples in the relation `\textit{citizen-of}'
are contained in the set of tuples in the relation `\textit{lives-in}' whose
second component is `\textit{Canada}'.
\end{itemize}
However, we require that the \emph{dependency graph} $G(D)$ of the ontology
to be a \emph{forest of trees}. The dependency graph is a labeled directed
graph that is constructed as follows: the nodes of the graph are the relational
symbols in the signature, and there is an edge $(R,P)$ in the graph iff
there is an rule of the form \eqref{eq:orule}. The edge is labeled with
the \emph{bindings} induced by the rule; e.g., if $\textit{descendant}(x,y)$
and $\textit{noble}(z)$ are two predicates in the signature and there
is the rule \eqref{eq:QE2}, then there is an edge from \textit{descendant}
to \textit{noble} labeled with the bindings $\{x=\textit{Elizabeth II},y=z\}$.

\subsection{Services and Mappings}

The available services are represented by means of another signature
$S=\tau=\tup{S_1^{s_1},\ldots,S_k^{s_k}}$ called the \emph{services signature},
where each symbol $S_i$ represents a concrete service in the Internet
that ``offers'' some information.

The semantic description of services is expressed following the
LAV paradigm in terms of mappings. These mappings describe the
services in terms of concepts in the domain ontology \cite{Ullman00}:
for each service $S_i$, there is a mapping that describes $S_i$
as a \emph{conjunctive query} on the concepts in the ontology that
also distinguish input and output attributes of the service.
For example, a service $S(x,y)$ that returns information about
flights originating at a given US city can be described with the view:
\[ S(\$x,y)\ \qrule\ \flight(x,y), \UScity(x)\,, \]
where $flight^2$ and $uscity^1$ are relational symbols in the
ontology. The symbol `\$' denotes that $x$ is an input attribute. 
The semantic interpretation of a mapping like this one enforce the following:
\begin{itemize}
\item the service represented by $S$ provides information in the
      form of tuples $(x,y)$,
\item the service is called with $x$ as input attribute and returns $(x,y)$,
\item each tuple $(x,y)$ returned by the service satisfies the rhs
      of the view; i.e., $\flight(x,y)$ and $\UScity(x)$, and
\item the views are not necessarily \emph{complete}; i.e., there may be
      other tuples $(x,y)$ that satisfy the rhs of the view but which
      are not available through $S$.
\end{itemize}

The LAV approach is commonly used in integration systems because it permits
the scalability of the system as new services become available \cite{Ullman00}.
Under LAV, the appearance of a new service only causes the addition of a new
mapping describing the service in terms of the concepts in the ontology.
Under GAV, on the other hand, the ontology concepts are semantically described
using views in terms of the sources of information.
Thus, the extension or modification of the ontology is an easy task in GAV as
it only involves the addition or local modification of few descriptions
\cite{Ullman00}.
Therefore, the LAV approach is best suited for applications with a stable
ontology but with changing data sources whereas the GAV approach is best
suited for applications with stable data sources and a changing ontology.
For the Semantic Web, we assume that the ontology of concepts is the stable
component. We believe that this is a reasonable assumption since, once a
common language is agreed upon to describe Web resources, the only changing
characteristic is the number and nature of resources which constantly pop
up or disappear from the Web.

Up to here, we have described all elements in the integration framework
$IS=\tup{D,S,M}$ where $D=\tup{\sigma,A}$ is an ontology of concepts, 
$S=\tau$ represent the available services in the Web and $M$ is a
collection of LAV mappings describing the services in terms of the
concepts in $D$.
The integration framework can be thought as the ``knowledge base'' (KB)
in a system designed for answering requests about the selection and
composition of Web services.
Ideally, the KB should support the efficient processing of user requests.

@@

A user request is a tuple $R=\tup{Q,P}$ where $Q$ is a conjunctive query
in terms of concepts in the ontology that describes how these concepts
must be combined to resolve a given task, and a set $P$ of preferences.
For example, the query:
\[ Q(x)\ \qrule\ \flight(\LA,x), \flight(x,\PA). \]
can be used to find all cities on which a two-leg flight
from Los Angeles to Paris stop. Continuing with the previous example,
this query can be answered using the view represented by $S$
as $I(x) \qrule S(\LA,x), S(x,\PA)$. This rewriting is correct yet
not necessarily complete as there may be two-leg flights from
Los Angeles to Paris that stop at non-US cities, e.g.,\ London.

The preferences are used to rank the collection of valid rewritings.
Once this ranking is obtained, the solution for the request $R$ is
any best-ranked and valid rewriting.
In this work, we consider a simple yet expressive language for 
preferences in which each preference is treated as a ``soft constraint''
on the rewriting of the query.
A soft constraint is a tuple $\pi=\tup{\varphi,c}$ where $\varphi$
is a propositional formula and $c$ is the cost associated with $\varphi$.
The idea is that each valid rewriting is associated with a cost equal to
the sum of the costs of the preferences violated by the rewriting,
and that the best ranked rewriting is one that has minimum cost among
all valid ones.

It only remains to say what type of propositional formulas $\varphi$
are allowed and when a preference is violated by a rewriting.
The set of propositions for constructing preferences is
$\L(IS)=\{ R : R\in\sigma \} \cup \{S : S\in\tau \}$ that corresponds
to the relational symbols either in the ontology signature or in the
services signature. Elements of $\L(IS)$ are propositional symbols that
should not be confused with their relational interpretation in $IS$;
indeed, if the reader is more comfortable, he may think on a different
symbol altogether such as $P_R$, $[R]$, or other.
The validity of a preference is defined with respect to the propositional
model $\M(I)$ (truth assignment for the symbols in $\L(IS)$) constructed
from a valid rewriting $I(\bar x)$:
$\M(I)$ is defined by $S=\textbf{true}$ iff the service $S$ appears in
$I(\bar x)$, for $S\in\tau$, or by $R=\textbf{true}$ iff the concept $R$
appears in the unique path from a concept $R'$ to the root in the dependency
graph $G(D)$ where $R'$ appears in a service $S(\bar y)$ used in $I(\bar x)$,
for $R\in\sigma$.
That is, the model makes true the service symbols used in $I(\bar x)$, or
the ontology symbols used in services in $I(\bar x)$, or the ontology symbols
that can be reached from the latter in the dependency graph $G(D)$.
For example, the rewriting $I(x)\qrule S(\LA,x),S(x,\PA)$ defines
the model $\M(I)=\{S=\textbf{true},\flight=\textbf{true},\UScity=\textbf{true}\}$.

A preference $\varphi$ holds in an answer $I(\bar x)$ iff $\M(I)\vDash\varphi$.
This simple language permits us to express interesting preferences such as:
\begin{itemize}
\item \textbf{Hard constraints:} any constraint $\pi=\tup{\varphi,\infty}$
is treated as a hard constraint that must be satisfied by every rewriting;
i.e., if the best rewriting has infinite cost, then it violates a hard constraint,
\item \textbf{QoS preferences:} this type of preferences can be used to
assign absolute quantities of reward/cost to single services as the one
used for integrated QoS parameters. For example, if each service $S_i$
is associated with a QoS reward of $r_i$, then the collection of preferences
$\pi_i=\tup{\neg S_i,-r_i}$ selects a valid rewriting with services that
have the highest combined QoS,
\item \textbf{Conditional preferences:} a user's preference of the type
`if service $S$ is used, then service $R$ should be used as well' can be
modeled with the hard constraint $S \Rightarrow R$,
\item \textbf{Preferences of the type at-least-one:} a user's preference
of the type that at least one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the hard constraint $S_1\lor\cdots\lor S_n$, and
\item \textbf{Preferences of the type at-most-one:} a user's preference
of the type that at most one of the services $S_1,\ldots,S_n$ should be
used in the rewriting, can be modeled with the collection
$\{\neg S_i\lor\neg S_j: i\neq j\}$ of hard constraints.
\end{itemize}

@@

We extend the \mcdsat\ system of Arvelo et al.\ \cite{arvelo:aaai06} for QRP.
An instance of QRP consists of a collection of views and a query on abstract
concepts.  The problem consists in rewriting the query in terms of the
views such that each tuple produced by the rewriting is a tuple of the
solution \cite{Ullman00}.
\mcdsat\ reduces QRP to the problem of finding the models of a propositional
logic theory that satisfies the following properties:
(1)~there is a 1-1 correspondence between the valid rewritings of the
query and the models of the logical theory,
(2)~given a model of the theory, one can recover the corresponding rewriting
in linear time, and
(3)~the theory can be constructed in polynomial time from the QRP instance.
Once the logical theory is constructed, one can be interested in finding all
minimal rewritings of the query as done in data integration systems with incomplete
sources, or just one rewriting as done when sources are complete \cite{Ullman00}.
For the former, off-the-shelf model enumeration tools such as c2d \cite{c2d} and
Relsat \cite{relsat} can be used, while off-the-shelf SAT solvers such as
Minisat \cite{minisat} or Rsat \cite{knot07lightweight} can be used in the latter case.

In this section, we have just enough space to explain how the logical
theory constructed by \mcdsat\ can be extended to capture the features
associated with SSPs that are not present in QRPs; namely, handling
constant symbols, input and output attributes, the ontology of concepts
with subsumption relationships, and user's preferences. The result is an
extended theory whose models are in correspondence with the valid solutions
of the SSP and, in the presence of preferences, whose \emph{best models}
are in correspondence with the best-ranked valid solutions of the SSP.

\subsubsection{Constant Symbols}
